Inferring true changes in microbial abundance from taxonomically-biased microbiome measurements

### 2022-01-19 conclusion.Rmd
Draft Conclusion

_This section is a placeholder for a short (2-3 paragraph) conclusion. We are planning to keep 'discussion' to the earlier sections and appendix, at least for initial submission._

>>>>

Our results show that even if taxonomic bias is consistent across samples at the taxonomic level being analyzed, it can create error in DA estimates that can lead to scientifically erroneous conclusions.
In particular, analyses based on fold-changes in proportions or in absolute abundances derived from total-abundance normalization (sometimes called 'Quantitative Microbiome Profiling') are both capable of leading to spurious results.
<!-- Perhaps state that we showed this theoretically, in mock communities, and in real case studies -->
The error, however, varies by experimental context.
It is negligible if the mean efficiency is stable across samples, and may be unimportant if it is not correlated with the focal covariate in a regression analysis.
The use of reference-taxon measurement and/or community calibration controls provides a way to validate and correct DA results.
In addition, bias-aware sensitivity and meta-analyses make it possible to rigorously account for bias even in the absence of such controls.
Over time, use of these methods will lead to a collective understanding of the experimental contexts in which these errors are likely.

We also showed that other DA methods are nearly or entirely invariant to consistent bias.
Analyses based on fold changes in ratios, including analysis of absolute abundance via reference-species normalization, are invariant to consistent bias in the MGS measurement.
In addition, careful pairing of total-abundance and MGS method may remove most of the bias-driven error.
Important questions remain about the true bias-invariance of these methods, including the extent to which bias is consistent across samples at the taxonomic resolution of the given MGS method and the interaction of bias with analytical assumptions such as how zeros are interpreted and the possibility of taxonomic misassignment and cross-sample contamination.
Nevertheless, these are straightforward, immediately implementable methods to addressing bias in fold-change analysis.

<!-- Placeholder for potential paragraph on limitations and open questions -->

In summary, our results show that protocol standardization is insufficient to nullify the effect of taxonomic bias in DA results, but that there are a range of practical solutions to mitigate it for an important subset of experimental questions and contexts.
The theoretical framework we present can be further used to rigorously analyze the effect of bias on other variations of DA analysis and of microbiome analyses more generally.
With this work, we look to a future in which microbiome researchers routinely choose an appropriate combination of experimental and data-analytic methods capable of answering their fundamental question while accounting for taxonomic bias and the other inherent limitations of MGS measurements---and in this way, gaining the confidence needed to codify statistical findings into true scientific knowledge.
